---
title: "MMAT Assignment1"
author: "Adil Hayat, Yijing Tao"
date: "2 September 2017"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rkdb)
```

1. _Our data table instinfo on the server has basic properties of all CME products.
Read enough of Q for Mortals 9.3 to issue simple queries, and answer the following
questions:_

    (a) _What are the products whose instrument codes are PA and ZC?_
        
        PA: Palladium(Precious Metals)
        
        ZC: Corn(Grain And Oilseed)
        
    (b) _If you see a trade in the market data for symbol RBU7 at price 15875, what was
the actual contract price? What is that price for gasoline in dollars per gallon?_
      
      The actual contract price will be $66675$. The price of gasoline per gallon was $1.5875
      
    (c) _If you buy one contract of RBU7 during the day at price 15875 according to
market data, and the settlement price on that day is 15876, how much money
does the clearing house deposit in your account?_
      
      The clearing house deposits $4.2 in my account.

2. _What were the 5 most actively traded interest rates products on August 18, 2017?
What were the 5 most actively traded non-interest rate products on the same date?_
```{r}
cat("Most Actively Traded by trade count:\n")
### most actively traded interest rate products
h <- open_connection('kx',6003)
query <- "select[5] from (`nT xdesc select nT:count i by sym 
          from trade where date=2017.08.18)"
data <- execute(h, query)
knitr::kable(data)

### most actively traded non-interest rate products
h <- open_connection('kx',7003)
query <- "select[5] from (`nT xdesc select nT:count i by sym
          from trade where date=2017.08.18)"
data <- execute(h, query)
knitr::kable(data)

cat("Most Actively Traded by Daily Volume:\n")
### most actively traded interest rate products
h <- open_connection('kx',6003)
query <- "select[5] from (`dailyVolume xdesc select dailyVolume:sum siz by sym
          from trade where date=2017.08.18)"
data <- execute(h, query)
knitr::kable(data)

### most actively traded non-interest rate products
h <- open_connection('kx',7003)
query <- "select[5] from (`dailyVolume xdesc select dailyVolume:sum siz by sym
          from trade where date=2017.08.18)"
data <- execute(h, query)
knitr::kable(data)

```


3. _Make a plot of traded volume on each date in August 2017 for the front-month 10-
Year Note Futures contract ZNU7. (There will be some interesting behavior in the last
week of this month, which we can see after I upload it next week.)_
```{r}
h <- open_connection('kx',6003)
query <- "select volume: sum siz by date from trade where sym=`ZNU7,
          date within 2017.08.01 2017.08.31"
data <- execute(h, query)
ggplot(data, aes(x=factor(format(date,"%d")), y=volume/1000000)) +geom_bar(stat="identity") + 
  ylab("Volume in Millions") + xlab("Date")
```



