---
title: "MMAT2"
author: "Adil Hayat, Yijing Tao"
date: "10 September 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rkdb)
theme_update(plot.title = element_text(hjust = 0.5))
```

## Question 1
```{r, cache=TRUE}
h <- open_connection('kx',7009)
query <- "select avvol: avg v by txntime from (select v:sum siz by date, 
          900000000000 xbar txntime from trade where date within 
          2017.07.31 2017.09.01, sym2inst[sym]=`ZW)"
data <- execute(h, query)

# create an empty data frame
df <- data.frame(txntime = seq(-25200,57600,900),dat=rep(0,93)) 
df <- merge(df,data,all.x = TRUE)
df$avvol <- ifelse(is.na(df$avvol),df$dat,df$avvol)

ggplot(df, aes(x=txntime/3600, y=avvol/1000)) +geom_line(color='red') + 
  geom_point() + ylab("Thousands of lots in 15 minutes") + xlab("Time") +
  ggtitle("Wheat(ZW) 2017-07-31 to 2017-09-01")+ 
  scale_x_continuous(limits=c(-7,16), breaks=seq(-6,16,2))
  
```
We think the spikes in the graph are caused at market open/close and at the times of release of key market news/numbers.


## plot the graph for various other products
```{r, cache=TRUE}
# interest rate products
h <- open_connection('kx',6009)
query <- "select first inst, first name by class from `class`v xdesc (select inst,
class,name from instinfo) lj select v:sum siz by inst:sym2inst[sym] from trade where
date within 2017.07.31 2017.09.01, not null sym2inst[sym]"
prod <- execute(h,query)


for(inst in prod$inst){
  h <- ifelse(inst=="ZN",open_connection('kx',6009),open_connection('kx',7009))
  query <- paste("select avvol: avg v by txntime from (select v:sum siz by date, 
                 900000000000 xbar txntime from trade where date within 
                 2017.07.31 2017.09.01, sym2inst[sym]=","`",inst,")",sep="")
data <- execute(h, query)

# for plotting
# create an empty data frame
df <- data.frame(txntime = seq(-25200,57600,900),dat=rep(0,93)) 
df <- merge(df,data,all.x = TRUE)
df$avvol <- ifelse(is.na(df$avvol),df$dat,df$avvol)

print(ggplot(df, aes(x=txntime/3600, y=avvol/1000)) +geom_line(color='red') + 
        geom_point() + ylab("Thousands of lots in 15 minutes") + xlab("Time") +
    ggtitle(paste(inst,"2017-07-31 to 2017-09-01"))+ 
  scale_x_continuous(limits=c(-7,16), breaks=seq(-6,16,2)))
}

```



## Question 2
```{r, cache=TRUE}
h <- open_connection('kx',6009)
query <- "update perOfTotal:v*100%sum(v) from update time2expir:expir-min(expir) from 
update expir: sym2expir[sym] from select v:sum siz by sym from trade where date 
within 2017.08.01 2017.09.01,sym2inst[sym]=`GE"
volData <- execute(h, query)
ggplot(volData, aes(x=time2expir, y=perOfTotal)) +geom_line(color='green') +
  geom_point() + ylab("Percentage of total volume") + xlab("Months to expiration") + 
  ggtitle("Eurodollar(GE) 2017-07-31 to 2017-09-01")

W = sum(-volData$perOfTotal/100*log(volData$perOfTotal/100))
Neff = exp(W)

cat("W=",W,"\n")
cat("Neff=",Neff,"\n")
```
## plot the graph for the 2 most active IR instruments
```{r, cache=TRUE}
# get 2 most active IR instruments
h<- open_connection("kx",6009)
query <- "2# select inst from `cumvol xdesc select cumvol:sum siz by 
          inst:sym2inst[sym] from trade where date within 
          2017.07.31 2017.09.01, not null sym2inst[sym]"
insts <- execute(h,query) 
insts$W <- rep(NA,nrow(insts))
insts$Neff <- rep(NA,nrow(insts))

for(i in 1:nrow(insts)){
  inst <- insts$inst[i]
  query <- paste("update perOfTotal:v*100%sum(v) from update time2expir:expir-min(expir)
    from update expir: sym2expir[sym] from select v:sum siz by sym from trade where date 
    within 2017.07.31 2017.09.01,sym2inst[sym]=","`",inst,sep = "")
volData <- execute(h, query)
print(ggplot(volData, aes(x=time2expir, y=perOfTotal)) +geom_line(color='green') +
        geom_point() + ylab("Percentage of total volume") + 
        xlab("Months to expiration") + ggtitle(paste(inst,"2017-07-31 to 2017-09-01")))

W = sum(-volData$perOfTotal/100*log(volData$perOfTotal/100))
Neff = exp(W)
insts$W[i] <- W
insts$Neff[i] <- Neff
}

knitr::kable(insts)
```

## plot the graph for 5 most active non-interest rate instruments
```{r, cache=TRUE}
# get 5 most active non-IR instruments
h<- open_connection("kx",7009)
query <- "5# select inst from `cumvol xdesc select cumvol:sum siz by 
          inst:sym2inst[sym] from trade where date within 2017.07.31 2017.09.01,
          not null sym2inst[sym]"
insts <- execute(h,query) 
insts$W <- rep(NA,nrow(insts))
insts$Neff <- rep(NA,nrow(insts))

for(i in 1:nrow(insts)){
  inst <- insts$inst[i]
  query <- paste("update perOfTotal:v*100%sum(v) from update time2expir:expir-min(expir) 
from update expir: sym2expir[sym] from select v:sum siz by sym from trade where date 
within 2017.07.31 2017.09.01,sym2inst[sym]=","`",inst,sep = "")
volData <- execute(h, query)
print(ggplot(volData, aes(x=time2expir, y=perOfTotal)) +geom_line(color='green') +
        geom_point() + ylab("Percentage of total volume") + 
        xlab("Months to expiration") + ggtitle(paste(inst,"2017-07-31 to 2017-09-01")))

W = sum(-volData$perOfTotal/100*log(volData$perOfTotal/100))
Neff = exp(W)
insts$W[i] <- W
insts$Neff[i] <- Neff
}

knitr::kable(insts)

```

EuroDollar(GE) has large $N_{eff}$, i.e. more than a few maturities active at a time. 

## Question 3
# functions to get eta and avq2t trade data
```{r, cache=TRUE}

getEta<- function(h, sym, date){
  arg <- paste("[",gsub("-",".",date),";","`",sym,"]", sep="")
  query<- paste("{[d;s] select eta:(sum 1-rev)%(sum rev) from update rev:differ up from 
                select up:chg>0 from select chg:deltas prc from 
                (select prc,diff:differ prc from trade where date=d,sym=s) where diff}",
                arg,sep="")
return(execute(h,query)$eta)
}

getAvQ2T<- function(h, sym, date){
    arg <- paste("[",gsub("-",".",date),";","`",sym,"]", sep="")
    query <- paste("{[d;s]select avq2t from update avq2t: (bszsum + aszsum)%tszsum from
                   select bszsum: sum bsiz, aszsum: sum asiz, tszsum: sum tradesiz from
                   select from aj[ `seq; select distinct seq,tradesiz:siz from trade where 
                   date=d, sym=s;select seq,bsiz,asiz from quote where date=d, sym=s] where
                   not null bsiz}", arg, sep="")
 return(execute(h, query)$avq2t)
}
```


```{r, cache=TRUE}
# select the most active instrument on a particular date for the date 
# range and calculate the parameters
# interest rate products
h <- open_connection('kx',6009)
query <- "{[range] select first sym by inst, date from `inst`cumvol 
          xdesc update inst:sym2inst[sym] from select cumvol:sum siz by sym,date 
          from trade where date within range, not null sym2inst[sym]}[2017.08.01 2017.08.31]" 
intProdDf <- execute(h, query)

# get the eta and avq2t for the most active symbol
intProdDf$eta <- apply(intProdDf,1,function(x) getEta(h,x['sym'],x['date']))
intProdDf$avq2t <- apply(intProdDf,1,function(x) getAvQ2T(h,x['sym'],x['date']))

# aggregate the data on per instrument basis
intPDf <- intProdDf %>% group_by(inst) %>% summarise(eta=mean(eta),
                                                     avq2t=mean(avq2t))

# non-interest rate products
h <- open_connection('kx',7009)
query <- "{[range] select first sym by inst, date from `inst`cumvol xdesc 
          update inst:sym2inst[sym] from select cumvol:sum siz by sym,date
          from trade where date within range, not null sym2inst[sym]}[2017.08.01 2017.08.31]" 
nonIntProdDf <- execute(h, query)

# get the eta and avq2t for the most active symbol
nonIntProdDf$eta <- apply(nonIntProdDf,1,function(x) getEta(h,x['sym'],x['date']))
nonIntProdDf$avq2t <- apply(nonIntProdDf,1,function(x) getAvQ2T(h,x['sym'],x['date']))

# aggregate the data on per instrument basis
nonIntPDf <- nonIntProdDf %>% group_by(inst) %>% summarise(eta=mean(eta),
                                                           avq2t=mean(avq2t))

```




```{r, cache=TRUE}
# plot the data
plotDf <- rbind(nonIntPDf,intPDf)
ggplot(plotDf,aes(x=avq2t, y=eta)) + geom_text(aes(label=inst), size=3) + 
  xlab("Average quote size/average aggresive trade size") + 
  ylab(expression("Reversion parameter "*eta)) +
  scale_x_continuous(trans="log", breaks=c(2,5,10,20,50,100,200,500,1000,2000)) + 
  scale_y_continuous(limits=c(0,1.5)) + ggtitle("2017-07-31 to 2017-09-01")
```

