---
title: "RMHW3"
author: "Adil Hayat"
date: "20 September 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(0)
theme_update(plot.title = element_text(hjust = 0.5))
```


## Problem 1(a)
We are given,

$$\lim_{x \to \infty} x^\alpha(1-F(x)) = b$$

\[
  \begin{aligned}
  P\bigg(\frac{M_n}{(bn)^{1/\alpha}}\bigg) &=P(M_n <= x(bn)^{1/\alpha})\\
  & = F^n(x(bn)^{1/\alpha})\\
  & = (1-(1-F(x(bn)^{1/\alpha})))\\
  & = \bigg(1 - \frac{x^\alpha(bn)}{x^\alpha(bn)}(1-F(x(bn)^{1/\alpha}))\bigg)\\
  & \approx \bigg(1-\frac{1}{x^{\alpha}n}\bigg)\\
  & \to  e^{-x^{-\alpha}}
  \end{aligned}
\]

So, $\{M_n/(bn)^{1/\alpha}\}^{\infty}_{n=1}$ converges weakly to a limiting Frechet distribution with cdf given by $e^{-x^{-\alpha}}$


## Problem 1(b)
```{r}
alpha <- 10
rand <- runif(1000)
frechet <- (1/-log(rand))^(1/alpha)

hist(frechet,10, freq=FALSE, main="Frechet Distribution")
```


## Problem 1(c)
\[
  \begin{aligned}
  F(x)&=1-\bigg(\frac{c}{c+x}\bigg)^{\alpha}\\
  U& = 1-\bigg(\frac{c}{c+x}\bigg)^{\alpha}\\
  \frac{c}{c+x}& = (1-u)^{1/\alpha}\\
  x& = \frac{c}{(1-u)^{1/\alpha}}-c\\
  \end{aligned}
\]

We try to find the corresponding limiting distribution
\[
  \begin{aligned}
  P\bigg(\frac{M_n-a_n}{b_n}\leq x\bigg) &= P(M_n \leq a_n + b_nx)\\
  & = F^n(a_n + b_n x)\\
  & = (1-(1-F(a_n + b_n x)))^n\\
  & = \bigg(1 - \frac{c^\alpha}{(c + a_n + b_n x)^\alpha}\bigg)^n\\
  &Substituting\hspace{0.1cm} a_n = -c \hspace{0.1cm} and \hspace{0.1cm} b_n = cn^{1/\alpha}\\
  & \approx \bigg(1-\frac{c^\alpha}{c^{\alpha}nx^\alpha}\bigg)^n\\
  & \approx \bigg(1-\frac{x^{-\alpha}}{n}\bigg)^n\\
  & \to  e^{-x^{-\alpha}}
  \end{aligned}
\]
 
 which is a Frechet distribution as in part 1(a)

Let's  choose $c = 1$ and $\alpha =10$ 
```{r, cache=TRUE}
alpha <- 10
c <- 1
N <- 1000

# generate the data from exact distribution
M <- rep(0,N)

for(i in 1:length(M)){
  M[i] <- (max(c/(1-runif(250))^(1/alpha) - c) + c)/(c*i^(1/alpha))
}

# generating the frechet distribution data
rand <- runif(N)
frechet <- (1/-log(rand))^(1/alpha)

# overlay the plots
plotDf <- data.frame(samples=c(frechet, M))
plotDf$dataSource <- c(rep("Frechet",N), rep("Theoretical",N))

ggplot(plotDf, aes(samples, fill = dataSource)) + geom_density(alpha = 0.2) + 
  ggtitle("Theoretical vs Estimated Frechet Distribution")
```



## Problem 2(a)
This is the case of polynomail Frechet distribution.
We need to find $\alpha$, b for which

$$x^{\alpha}(1-F(x)) \to b \hspace{1cm} as \hspace{0.2cm} x \to \infty$$
\[
  \begin{aligned}
  x^{\alpha}\int^{\infty}_{x}f_4(u)du & =  \frac{\int^{\infty}_{x}f_4(u)du}{x^{-\alpha}}\\
  \end{aligned}
\]

Using L'Hospital Rule, 

$$\frac{-f_4(x)}{-\alpha x^{-(\alpha+1)}} \to b  \hspace{0.2cm }as \hspace{0.2cm } x\to \infty$$
Substituting the value of $f_4(x)$, we get 

\[
  \begin{aligned}
  f_4(x) &= \frac{\Gamma\left(5/2\right)}{\sqrt{4\pi}\Gamma\left(2\right)}{\bigg(1+\frac{x^2}{4}\bigg)}^{-5/2}\\
   & =  \frac{3}{8}{\bigg(1+\frac{x^2}{4}\bigg)}^{-5/2}\\\\
  \end{aligned}
\]

So, we can write the limit as,

\[
  \begin{aligned}
  \lim_{x \to \infty}\frac{\frac{3}{8}{\bigg(1+ \frac{x^2}{4}\bigg)}^{-5/2}}{\alpha x^{-(\alpha+1)}}&=b\\
   \lim_{x \to \infty}\frac{3}{8\alpha}\frac{1}{ {\bigg(1+ \frac{x^2}{4}\bigg)}^{5/2}x^{-(\alpha+1)}}&=b\\\\
  \end{aligned}
\]

Setting $\alpha=4$, the equation reduces to
\[
  \begin{aligned}
   \lim_{x \to \infty}\frac{3}{8*4}\frac{1}{{\bigg(x^2+ \frac{1}{4}\bigg)}^{5/2}}& =b\\
   \frac{3}{8*4}\frac{1}{\frac{1}{4}^{5/2}}& = b\\
   b & =3\\
  \end{aligned}
\]



## Problem 2(b)
```{r}
N <- 10000
df <- 4
rand <- -rt(N,df)
dat <- sort(rand[rand>0],decreasing = TRUE)
plotDat <- rep(0,length(dat)-1)
for(i in 1:length(dat)-1){
  plotDat[i] <- 1/(sum(log(dat[1:(i+1)]))/(i+1) - log(dat[i+1]))   
}

plotDfHill <- data.frame(dat=plotDat)
plotDfHill$seq <- seq(2,length(dat),1)
ggplot(plotDfHill[1:500,], aes(y=dat, x=seq)) + geom_line() + xlab("k") + 
  ylab(expression(alpha))+geom_hline(yintercept = 3.6, color="red") +
  annotate("text", x=200, y=4.5, label="alpha=3.6") +
  ggtitle("Hill Plot for t-4 Distribution")

```

From the plot we observe that $\alpha=3.6$ which is close to the theoretical value of $\alpha=4$. 

## Problem 2(c)
```{r}
# plot the POT plot
dat <- sort(rand[rand>0])
e <- rep(0,length(dat))

for(i in 1:length(dat)){
  e[i] <- sum(pmax(dat[(i+1):length(dat)]-dat[i],0))/length(dat[dat>dat[i]])
}

plotDf <- data.frame(y=e)
plotDf$x <- dat

ggplot(plotDf, aes(y=y, x=x)) + geom_line() +
  xlab("Losses") + ggtitle("POT plot for t-4 distribution")+ ylab("e(u)")

```

From the plot we observe that the $\theta=0.2$.

## Problem 3
```{r, cache=TRUE}
df <- read.csv("ibm091716.csv")

# plot the hill plot
ret <- -df$return
dat <- sort(ret[ret>0],decreasing = TRUE)
plotDat <- rep(0,length(dat)-1)
for(i in 1:length(dat)-1){
  plotDat[i] <- 1/(sum(log(dat[1:(i+1)]))/(i+1) - log(dat[i+1]))   
}

plotDfHill <- data.frame(dat=plotDat)
plotDfHill$seq <- seq(2,length(dat),1)
ggplot(plotDfHill[1:500,], aes(y=dat, x=seq)) + geom_line() + xlab("k") + 
  ylab(expression(alpha)) + geom_hline(yintercept = 3.4, color="red")

# plot the POT plot
dat <- sort(ret[ret>0])
e <- rep(0,length(dat))

for(i in 1:length(dat)){
  e[i] <- sum(pmax(dat[(i+1):length(dat)]-dat[i],0))/length(dat[dat>dat[i]])
}

plotDf <- data.frame(y=e)
plotDf$x <- dat

ggplot(plotDf, aes(y=y, x=x)) + geom_line() + geom_abline(intercept=0,slope=0.4, color="red") +
  xlab("Losses") + ylab("e(u)")


```

Observing the graphs,
The value of $\alpha$=3.4.
This gives us the value of $\theta$= 0.28


## Problem 4
Option (b) is incorrect as EVT highlights the distributions justified by central limit theorem cannot be used for extreme value estimation.


## Problem 5
Option (d) is the only incorrect option as we do not ignore extreme events.


## Problem 6
The information ration $IR=$ `r (7.8-7.2)/1.25`. The number of years for which the performance should persist can be derived as:
$\frac{0.6T}{1.25\sqrt{T}}>=1.96$ which implies that T>=16.7. So, option (a) is correct.


## Problem 7
Assuming the risk-free rate to be 5% and filling the values in the table.
\begin{table}[ht!]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline
        Fund & Expected Return & Volatility & Absolute Risk & Information Ratio & Exp. Residual Ret. & Sharpe Ratio \\\hline\hline
        A & 8.4\% & 14.3\% & 1.82\% & 1.1 &  2\%    & 0.237\\\hline
        B & 8.56 & 16.4\% & 2.4\% & 0.9 &  2.16\%  &    0.217\\\hline
        C & 8.35 & 17.8\% & 1.5\% & 1.3 &  1.95\% & 0.188\\\hline
        D & 8.5\% & 19.1\% & 1.8\% & 1.17 & 2.1\% & 0.183\\\hline
    \end{tabular}
    \caption{Fund Performance}
\end{table}

Using the values in the table above, we see that only fund A and B satisfy the guidelines (I) and (II) with Fund D having the highest information ratio, So, Fund A should be chosen. 


